---
title: "Is TWiV getting longer?"
author: "Neil Saunders"
date: "compiled `r Sys.time()`"
output: 
  github_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = FALSE, message = FALSE, warning = FALSE, fig.path = "../../figures/")
library(tidyverse)
library(lubridate)
library(rvest)
library(xml2)
library(ggridges)
library(tidytext)
library(wordcloud)
library(knitr)

theme_set(theme_bw())

# function to make data frame given podcast RSS URL
feed_to_df <- function(rss) {
  feed_items <- read_xml(rss) %>%
  xml_nodes("item")
  
  feed_df <- data.frame(pubDate = feed_items %>% xml_nodes("pubDate") %>% xml_text(),
                        title = feed_items %>% xml_nodes("title") %>% xml_text(),
                        encLength = feed_items %>% xml_nodes("enclosure") %>% xml_attr("length"),
                        duration = feed_items %>% xml_nodes("itunes\\:duration") %>% xml_text(),
                        stringsAsFactors = FALSE) %>%
    mutate(pubDate = as.POSIXct(strptime(pubDate, "%a, %d %b %Y %H:%M:%S +0000"))) %>%
    mutate(encLength = as.numeric(encLength),
           duration = ifelse(grepl(":\\d+:", duration), duration, paste0("00:", duration)),
           duration_seconds  = as.numeric(hms(duration))
           )
  feed_df
}
```

# Introduction
Is science podcast [This Week In Virology](http://www.microbe.tv/twiv/) getting longer? And some other TWiV trivia.

# Get the XML
Data for every podcast episode is contained in [the RSS feed](http://twiv.microbeworld.libsynpro.com/twiv). RSS is XML, so reading it into R is very easy. Details for each episode are found in the `item` nodes.

# Make the data frame
There is sure to be a clever and elegant way to map straight into a data frame using _e.g._ `purrr`. This is not that way.

- `pubDate` is converted to date-time
- `enclosure length` (bytes) is converted to numeric values
- `duration` is pre-pended with "00" where hours are absent, then converted to seconds as numeric values

```{r parse-data}
twiv_df <- feed_to_df("http://twiv.microbeworld.libsynpro.com/twiv")
```

The end result is a data frame with rows that look like this:

```{r-df-rows}
twiv_df %>%
  top_n(5, wt = pubDate) %>%
  kable()
```

# Is TWiV getting longer?
Some time in 2016, I started to form the impression that most episodes of TWiV were around the two hour mark or more. Unfortunately, this prompted my decision to stop listening since I listen to several other podcasts and could not justify 2 hours each week for just one of them.

## Scatter plot
Let's start with a scatter plot.

There's a clear increase in duration for the first 4 years of the podcast, which then levels off. There's also quite a lot of noise due to "outliers": special episodes that tend to be much shorter than average.

```{r twiv-scatter}
twiv_df %>% 
  mutate(duration_minutes = duration_seconds/60) %>% 
  ggplot(aes(pubDate, duration_minutes)) + 
    geom_point() + 
    geom_smooth() + 
    labs(x = "Date", y = "Duration (minutes)", title = "This Week in Virology Episode Duration") + 
    scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")
```

## Summarise by year
How about if we plot the mean episode duration by year?

Mean duration passed 90 minutes in 2011 and approached 100 minutes in 2015. There is a very slight decrease in the mean duration in the episodes aired so far for 2017.

```{r twiv-bar}
twiv_df %>% 
  mutate(duration_minutes = duration_seconds/60,
         Year = year(pubDate)) %>%
  group_by(Year) %>%
  summarise(mean_duration = mean(duration_minutes, na.rm = TRUE)) %>%
  ggplot(aes(Year, mean_duration)) + 
    geom_col(fill = "skyblue3") + 
    labs(x = "Year", y = "Mean duration (minutes)", title = "This Week in Virology Mean Episode Duration By Year") +
<<<<<<< HEAD
    scale_x_continuous(breaks = 2008:year(Sys.Date()))
=======
    scale_x_continuous(breaks = 2008:2020)
>>>>>>> 0ed96ed2937f99411007b5710993c6e3f2527e01
```

## Distribution of duration
The "ridge line" plot shows duration distributions by year.

This clearly shows the duration creeping towards two hours in 2016, but dropping back a little in 2017.

```{r twiv-joy}
twiv_df %>% 
  mutate(duration_minutes = duration_seconds/60,
<<<<<<< HEAD
         Year = factor(year(pubDate), levels = year(Sys.Date()):2008)) %>%
=======
         Year = factor(year(pubDate), levels = 2020:2008)) %>%
>>>>>>> 0ed96ed2937f99411007b5710993c6e3f2527e01
  filter(!is.na(duration_minutes)) %>%
  ggplot(aes(x = duration_minutes, y = Year)) + 
    geom_density_ridges() + 
    theme_ridges() + 
    labs(x = "Duration (minutes)", y = "Year", title = "This Week in Virology Episode Duration Distribution By Year") +
    scale_x_continuous(breaks = seq(0, 150, 30), limits = c(0, 150))
```

# Conclusions
TWiV duration moved to around 90 minutes on average in 2010 - 2011, then moved again closer to 100 - 120 minutes in 2015 - 2016. However, the duration seems to have dropped back a little in 2017.

How about some of the other Microbe TV podcasts?

## This Week in Microbiology (TWiM)
Let's try comparing these in the same plot.
```{r twim-joy}
twim_df <- feed_to_df("http://feeds.feedburner.com/twim?format=xml")

twiv_df %>% 
  mutate(show = "This Week in Virology") %>% 
  bind_rows(mutate(twim_df, show = "This Week in Microbiology")) %>% 
<<<<<<< HEAD
  mutate(duration_minutes = duration_seconds/60, Year = factor(year(pubDate), levels = year(Sys.Date()):2008)) %>% 
=======
  mutate(duration_minutes = duration_seconds/60, Year = factor(year(pubDate), levels = 2020:2008)) %>% 
>>>>>>> 0ed96ed2937f99411007b5710993c6e3f2527e01
  filter(!is.na(duration_minutes)) %>% 
  ggplot(aes(x = duration_minutes, y = Year)) + 
    geom_density_ridges(aes(fill = show)) + 
    theme_ridges() + 
    labs(x = "Duration (minutes)", y = "Year", title = "Episode Duration Distribution by Year") + 
    scale_x_continuous(breaks = seq(0, 150, 30)) + 
  scale_fill_manual(values = c("darkorange", "skyblue3"))
```

Side by side might be better?
```{r twim-joy2}
twiv_df %>% 
  mutate(show = "This Week in Virology") %>% 
  bind_rows(mutate(twim_df, show = "This Week in Microbiology")) %>% 
  mutate(duration_minutes = duration_seconds/60, 
<<<<<<< HEAD
         Year = factor(year(pubDate), levels = year(Sys.Date()):2008), 
=======
         Year = factor(year(pubDate), levels = 2020:2008), 
>>>>>>> 0ed96ed2937f99411007b5710993c6e3f2527e01
         show = factor(show, levels = c("This Week in Virology", "This Week in Microbiology"))) %>% 
  filter(!is.na(duration_minutes)) %>% 
  ggplot(aes(x = duration_minutes, y = Year)) + 
    geom_density_ridges() + 
    theme_ridges() + 
    labs(x = "Duration (minutes)", y = "Year", title = "Episode Duration Distribution by Year") + 
    scale_x_continuous(breaks = seq(0, 150, 30)) + 
  facet_wrap(~show)
```

## This Week in Parasitism (TWiP)
We'll try these side by side.
```{r twip-joy}
twip_df <- feed_to_df("http://twip.microbeworld.libsynpro.com/rss")

twiv_df %>% 
  mutate(show = "This Week in Virology") %>% 
  bind_rows(mutate(twip_df, show = "This Week in Parasitology")) %>% 
  mutate(duration_minutes = duration_seconds/60, 
<<<<<<< HEAD
         Year = factor(year(pubDate), levels = year(Sys.Date()):2008), 
=======
         Year = factor(year(pubDate), levels = 2020:2008), 
>>>>>>> 0ed96ed2937f99411007b5710993c6e3f2527e01
         show = factor(show, levels = c("This Week in Virology", "This Week in Parasitology"))) %>% 
  filter(!is.na(duration_minutes)) %>% 
  ggplot(aes(x = duration_minutes, y = Year)) + 
    geom_density_ridges() + 
    theme_ridges() + 
    labs(x = "Duration (minutes)", y = "Year", title = "Episode Duration Distribution by Year") + 
    scale_x_continuous(breaks = seq(0, 150, 30)) + 
  facet_wrap(~show)
```

# Other ways to analyse the feed data
## Keywords
A word cloud of the podcast keywords. We remove stop words plus the words "virus", "viruses", "virology", "viral", "twiv".

Keywords occurring 10 or more times:
```{r twiv-wordcloud}
data.frame(keywords = read_xml("http://twiv.microbeworld.libsynpro.com/twiv") %>% xml_nodes("item itunes\\:keywords") %>% xml_text(), 
           stringsAsFactors = FALSE) %>% 
  unnest_tokens(words, keywords) %>% 
  count(words) %>% 
  filter(!words %in% c("virus", "viruses", "virology", "viral", "twiv", stop_words$word)) %>% 
  with(wordcloud(words, n, min.freq = 10, random.order = FALSE, colors = brewer.pal(11, "Spectral")))
```